---
title: "Splitting–coalescence–estimation method (SCEM)"
author: "Joe Roe"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The splitting–coelescene–estimation method (SCEM) is an algorithm for estimating the birth season of animals from oxygen stable isotope time series data.
It was developed by Chazin et al. [(2018)](https://doi.org/10.1111/arcm.12432) to improve on the cosine method previously developed by Tornero et al. [(2013)](https://doi.org/10.1016/j.jas.2013.05.013).

This vignette is derived from the [usage example](https://github.com/soudeepd/SCEM/blob/master/Usage-Example.R) included with Chazin et al.'s original R code. 
It has been adapted to follow a 'tidy' workflow, but otherwise replicates the analysis in their 2018 *Archaeometry* paper.

## Setup and read data

We'll need this package and the `tidyverse` collection.

```{r, message=FALSE}
library(sits)
library(tidyverse)
```

We'll use the oxygen isotope from Tsaghkovit (Armenia) included in this package.
Subsequent analyses will be performed on each specimen individually, so let's group by specimen ID now.

```{r}
data("tsaghkahovit")
tsaghkahovit %>% 
  as_tibble() %>% 
  group_by(id) ->
  tsaghkahovit
```
```{r, echo=FALSE}
knitr::kable(tsaghkahovit[1:10,], align = c("l", "l", "l", "r"),
             caption = "First 10 rows of the Tsaghkahovit data")
```

## Cosine method

The cosine method is a simple approach to estimating season of birth, but is sensitive to the choice of original parameters. 

---

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
